---
title: "Constraint"
author: "Brendan Knapp"
output: 
  github_document:
    toc: true
editor_options: 
  chunk_output_type: console
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align = "center",
                      comment = "#>")
options(width = 200)
```

```{r}
library(igraph, warn.conflicts = FALSE)
```

```{r}
strike_net <- "*Vertices      24\r\n     1 \"Xavier\"    0.8220    0.7997\r\n     2 \"Utrecht\"    0.7189    0.5679\r\n     3 \"Frank\"    0.1500    0.8500\r\n     4 \"Domingo\"    0.3480    0.1500\r\n     5 \"Norm\"    0.5938    0.5446\r\n     6 \"Hal\"    0.3316    0.6461\r\n     7 \"Russ\"    0.8359    0.4228\r\n     8 \"Karl\"    0.4469    0.8339\r\n     9 \"Bob\"    0.4608    0.5138\r\n     10 \"Quint\"    0.8500    0.4964\r\n     11 \"Wendle\"    0.8001    0.8384\r\n     12 \"Ozzie\"    0.5793    0.7488\r\n     13 \"Ike\"    0.3135    0.5410\r\n     14 \"Ted\"    0.8127    0.2755\r\n     15 \"Sam\"    0.7070    0.6662\r\n     16 \"Vern\"    0.6746    0.3554\r\n     17 \"Gill\"    0.2525    0.7417\r\n     18 \"Lanny\"    0.4151    0.6905\r\n     19 \"Mike\"    0.3520    0.4907\r\n     20 \"Carlos\"    0.2922    0.1971\r\n     21 \"Alejandro\"    0.3967    0.3307\r\n     22 \"Paul\"    0.6923    0.4682\r\n     23 \"Eduardo\"    0.3200    0.1712\r\n     24 \"John\"    0.3770    0.6861\r\n*Arcs\r\n      1     11    1.0000\r\n      1     15    1.0000\r\n      2      5    1.0000\r\n      2      7    1.0000\r\n      2     10    1.0000\r\n      2     15    1.0000\r\n      3     17    1.0000\r\n      4     20    1.0000\r\n      4     21    1.0000\r\n      4     23    1.0000\r\n      5      2    1.0000\r\n      5      9    1.0000\r\n      5     12    1.0000\r\n      5     15    1.0000\r\n      5     16    1.0000\r\n      5     22    1.0000\r\n      6      9    1.0000\r\n      6     17    1.0000\r\n      6     24    1.0000\r\n      7      2    1.0000\r\n      7     10    1.0000\r\n      7     14    1.0000\r\n      8     12    1.0000\r\n      8     18    1.0000\r\n      8     24    1.0000\r\n      9      5    1.0000\r\n      9      6    1.0000\r\n      9     13    1.0000\r\n      9     18    1.0000\r\n      9     19    1.0000\r\n      9     21    1.0000\r\n      9     24    1.0000\r\n     10      2    1.0000\r\n     10      7    1.0000\r\n     10     22    1.0000\r\n     11      1    1.0000\r\n     11     15    1.0000\r\n     12      5    1.0000\r\n     12      8    1.0000\r\n     13      9    1.0000\r\n     13     17    1.0000\r\n     13     19    1.0000\r\n     14      7    1.0000\r\n     14     16    1.0000\r\n     15      1    1.0000\r\n     15      2    1.0000\r\n     15      5    1.0000\r\n     15     11    1.0000\r\n     16      5    1.0000\r\n     16     14    1.0000\r\n     17      3    1.0000\r\n     17      6    1.0000\r\n     17     13    1.0000\r\n     17     24    1.0000\r\n     18      8    1.0000\r\n     18      9    1.0000\r\n     18     24    1.0000\r\n     19      9    1.0000\r\n     19     13    1.0000\r\n     20      4    1.0000\r\n     20     21    1.0000\r\n     20     23    1.0000\r\n     21      4    1.0000\r\n     21      9    1.0000\r\n     21     20    1.0000\r\n     21     23    1.0000\r\n     22      5    1.0000\r\n     22     10    1.0000\r\n     23      4    1.0000\r\n     23     20    1.0000\r\n     23     21    1.0000\r\n     24      6    1.0000\r\n     24      8    1.0000\r\n     24      9    1.0000\r\n     24     17    1.0000\r\n     24     18    1.0000\r\n"
temp_file <- tempfile(fileext = ".net")
readr::write_file(strike_net, temp_file)

strike_g <- as.undirected(
  read_graph(temp_file, format = "pajek")
)

plot_igraph <- function(g, ...) {
  plot(g, asp = 0.5, 
       vertex.size = 4, vertex.label.dist = 1.5, ...)
}

plot_igraph(strike_g, main = "Strike")
```


# Ego Model

## UCINET Results

```
                   1
             Constra
                 int
             -------
 1    Xavier   1.125
 2   Utrecht   0.563
 3     Frank       1 ???? missing in ego.txt  
 4   Domingo   0.926
 5      Norm   0.236
 6       Hal   0.840
 7      Russ   0.611
 8      Karl   0.611
 9       Bob   0.287
10     Quint   0.611
11    Wendle   1.125
12     Ozzie   0.500
13       Ike   0.611
14       Ted   0.500
15       Sam   0.563
16      Vern   0.500
17      Gill   0.406
18     Lanny   0.840
19      Mike   1.125
20    Carlos   0.926
21 Alejandro   0.583
22      Paul   0.500
23   Eduardo   0.926
24      John   0.522
```

```{r}
ucinet_ego_results <- c(
  1.125, 0.563, 1, 0.926, 0.236, 0.840, 0.611, 0.611, 0.287, 0.611, 1.125, 
  0.500, 0.611, 0.500, 0.563, 0.500, 0.406, 0.840, 1.125, 0.926, 0.583, 0.500, 0.926, 0.522
)
```

## `{igraph}` Replication

Extract each vertex's ego (order = 1) network and run `igraph::constraint()` on them individually.

```{r}
ego_constraint <- function(g, .order = 1L, .round = 3L, .nm = FALSE) {
  if (!is_named(g)) {
    vertex_attr(g, "name") <- sprintf("n%d", seq_len(vcount(g)))
  }
  
  ego_nets <- make_ego_graph(g, order = .order) # get all ego networks
  ego_names <- vertex_attr(g, "name") # grab the egos' names
  
  # walk through each ego network and measure constraint of the ego vertex
  out <- mapply(function(.ego_net, .ego_name) { 
    constraint(.ego_net, nodes = .ego_name)
  }, ego_nets, ego_names)
  
  out <- round(out, .round)
  if (.nm) out else unname(out)
}
```

```{r}
ego_constraint(strike_g, .nm = TRUE)
```

## Confirmation

```{r}
data.frame(
  name = vertex_attr(strike_g, "name"), 
  `UCINET "Ego"` = round(ucinet_ego_results, 3),
  `ego_constraint()` = ego_constraint(strike_g),
  check.names = FALSE
)
stopifnot(
  identical(
    round(ucinet_ego_results, 2), 
    ego_constraint(strike_g, .round = 2)
  )
)
```


# "Whole Network"

Identical to `igraph::constraint()`.

## UCINET Results

```
                  3
              Const
              raint
              -----
 1    Xavier  0.953
 2   Utrecht  0.405
 3     Frank      1
 4   Domingo  0.866
 5      Norm  0.198
 6       Hal  0.536
 7      Russ  0.482
 8      Karl  0.469
 9       Bob  0.238
10     Quint  0.482
11    Wendle  0.953
12     Ozzie  0.500
13       Ike  0.506
14       Ted  0.500
15       Sam  0.464
16      Vern  0.500
17      Gill  0.326
18     Lanny  0.562
19      Mike  0.771
20    Carlos  0.866
21 Alejandro  0.583
22      Paul  0.500
23   Eduardo  0.866
24      John  0.418
```

```{r}
ucinet_whole_results <- c( 
  0.953, 0.405, 1, 0.866, 0.198, 0.536, 0.482, 0.469, 0.238, 0.482, 0.953,
  0.500, 0.506, 0.500, 0.464, 0.500, 0.326, 0.562, 0.771, 0.866, 0.583, 
  0.500, 0.866, 0.418
)
```

## `{igraph}` Replication

```{r}
constraint(strike_g)
```

## Confirmation

```{r}
stopifnot(
  identical(
    ucinet_whole_results,
    constraint(strike_g) %>% round(3L) %>% unname()
  )
)
```

```{r}
data.frame(
  name = vertex_attr(strike_g, "name"),
  `UCINET "Whole Network"` = ucinet_whole_results,
  `igraph::constraint()` = round(constraint(strike_g), 3L),
  check.names = FALSE,
  row.names = NULL
)
```

## "Whole Network"?

What UCINET is calling "Whole Network" is misleading.

```{r}
ego_g <- make_ego_graph(strike_g, nodes = "Xavier")[[1L]]
plot_igraph(ego_g, main = "Xavier's Ego")
```


```{r}
extended_ego_g <- make_ego_graph(strike_g, order = 2L, nodes = "Xavier")[[1L]]
plot_igraph(extended_ego_g, main = "Xavier's Extended Ego")
```

### Extended Ego

However UCINET calulates it, "Whole Network" and `igraph::constraint()` measure the "extended" ego networks (order = 2).


```{r}
extended_ego_constraint <- function(g, ...) {
  ego_constraint(g, .order = 2L, ...)
}

extended_ego_constraint(strike_g)
```

#### Confirmation

```{r}
stopifnot(
  identical(
    constraint(strike_g) %>% round(3) %>% unname(),
    extended_ego_constraint(strike_g)
  )
)
stopifnot(
  identical(
    extended_ego_constraint(strike_g), 
    ucinet_whole_results
  )
)
```

"Whole Network" may describe UCINET's other "Whole Network" Structural Holes metrics, but not constraint.

```{r}
data.frame(
  `igraph::constraint()` = round(constraint(strike_g), 3L),
  `extended_ego_constraint()` = extended_ego_constraint(strike_g),
  `UCINET "Whole Network"` = ucinet_whole_results,
  check.names = FALSE,
  row.names = NULL
)
```




# Maximum Constraint

Either interpretation of Burt's Constraint can yield values > 1.

```{r}
el <- matrix(
  c(1, 2, 
    1, 3, 
    2, 3), ncol = 2L, byrow = TRUE)

g <- graph_from_edgelist(el, directed = FALSE)
vertex_attr(g, "name") <- letters[seq_len(vcount(g))]
plot(g)
```

```{r}
data.frame(
    name = vertex_attr(g, "name"),
    Ego = ego_constraint(g),
    `"Whole Network"/Extended Ego` = constraint(g),
    check.names = FALSE,
    row.names = NULL
)
```


```{r}
el <- matrix(
  c(1, 2, 
    1, 3, 
    2, 3,
    3, 4), ncol = 2L, byrow = TRUE)

g <- graph_from_edgelist(el, directed = FALSE) 
vertex_attr(g, "name") <- letters[seq_len(vcount(g))]
plot(g)
```

```{r}
data.frame(
    name = vertex_attr(g, "name"),
    Ego = ego_constraint(g),
    `"Whole Network"/Extended Ego` = constraint(g),
    check.names = FALSE,
    row.names = NULL
)
```

```{r}
el <- matrix(
  c(1, 2, 
    1, 3, 
    2, 3,
    3, 4,
    4, 5), ncol = 2L, byrow = TRUE)

g <- graph_from_edgelist(el, directed = FALSE) 
vertex_attr(g, "name") <- letters[seq_len(vcount(g))]
plot(g)
```

```{r}
data.frame(
    name = vertex_attr(g, "name"),
    Ego = ego_constraint(g),
    `"Whole Network"/Extended Ego` = constraint(g),
    check.names = FALSE,
    row.names = NULL
)
```


